Data Provenance Tracking for Concurrent Programs

### 2015-02-04 paper.tex
pre-merge commit
latest latest

\newcount\draft\draft=0 % set to 0 for submission or publication
\newcount\cameraready\cameraready=0

\documentclass[pageno,nohyperref]{jpaper}

\ifnum\draft=1
  \input{revision}
  \usepackage{drafthead}
\fi
\usepackage{xxx}

\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\graphicspath{{./figs/} {./plots/}}

\usepackage[normalem]{ulem}

@@

\usepackage{xspace}
\newcommand{\addtodo}[1]{\textcolor{red}{[To do: #1]}}
\newcommand{\ctrap}{CTrap\xspace}
\newcommand{\ctraps}{CTraps\xspace}
\newcommand{\ctrapsfull}{CTraps-Full\xspace}
\newcommand{\ctrapsmm}{CTraps-NRR\xspace}
\newcommand{\lws}{LWS\xspace}
\newcommand{\lwt}{LWT\xspace}

\newcommand{\Caption}[1]{\begin{minipage}{.95\columnwidth} \caption{#1} \end{minipage} \vspace{-1.2ex}}
\newcommand{\CaptionWide}[1]{\caption{#1}\vspace{-2ex}}
%\doublespacing
%\authorinfo{Anonymous for Submission}{}

%\authorinfo{Brandon Lucia \and Luis Ceze}{University of Washington, Department of Computer Science and Engineering}{\{blucia0a,luisceze\}@cs.washington.edu\\http://sampa.cs.washington.edu\\{\bf In Submission --- Please Do Not Distribute}}

@@

@@

\thispagestyle{empty}
\begin{abstract}
%Monitoring inter-thread memory dependences in production is useful.  Doing so
%improves debugging information and provides information to dynamic analyses and
%runtime systems.  Prior work that continuously monitors cross-thread
%dependences typically has overheads too high for production.  
We propose {\em \lws} (\lws), a mechanism for tracking {\em data
provenance information} in multithreaded code in a production setting.  Last
writer slices dynamically track provenance of values by recording the thread
and operation that last wrote each variable. We show that this information
complements core dumps and greatly improves debugability. We also propose {\em
communication traps} (\ctraps), an extensible framework that uses \lws to interpose on operations that lead to inter-thread communication. We
show how to use \ctraps to implement multiple dynamic analysis tools for
mulithreaded programs. Our evaluation on server programs and PARSEC benchmarks
shows that \lws has low run time overheads (0--15\%) for many
applications,including  memcached, LevelDB, MySQL, and Apache. Our debugging
case studies of real, buggy programs shows that \lws are indeed
useful for debugging.  
\end{abstract}

@@

Our approach is to provide the programmer with data {\em provenance}
information that can be examined alongside the program's memory state ({\em
i.e.}, core dump).  Provenance information describes why a memory location
contains its value.  In this work we develop a new provenance tracking
mechanism called {\em \lws} (LWS).  A memory location's last writer slice is a
record containing the program point and the identifier of the thread that last
wrote data to that memory location.  As the program executes in production, we
collect last writer slices for each potentially shared memory location,
recording the provenance of the data in each location.  Last writer slices are
saved with a core dump and are available in the debugger.

@@

Motivated by these {\em ad hoc} efforts, we develop a dynamic analysis
framework called {\em Communication Traps} or {\em \ctraps}.  \ctraps uses the
provenance information provided by last writer slices to identify memory
accesses that lead to {\em communication} between the threads. Communication
occurs when a thread accesses memory last written by a different thread and
\ctraps delivers a software {\em trap} to a thread when it communicates with
another thread, giving it information about how communication occurred.
Programmer-defined {\em \ctraps Applications} can implement arbitrary analysis
using that information. By default, \ctraps has low overheads, appropriate for
production in many cases.  As our evaluation shows, the overhead of \ctraps
applications scales with the complexity of their analysis.
Figure~\ref{fig:overview} illustrates the relationship between last writer
slices, \ctraps, and \ctraps applications.

@@

\paragraph{Contributions of This Work}
To summarize, this work makes several contributions:
\begin{itemize}

\item{{\bf Last writer slicing} ({\em LWS}): a provenance tracking mechanism
useful for concurrency debugging.  We present an \lws implementation
efficient enough for production use.}

\item{{\bf Communication traps} (\ctraps): an extensible
framework for implementing concurrency analyses, built on \lws.  }

\item{{\bf Efficacy evaluation}: We show that \lws is useful for debugging with
a study of several real concurrency bugs.  We show \ctraps is useful for
building concurrency analyses by implementing two analyses from prior
work~\cite{cci,defuse,recon}. }

\item{{\bf Performance evaluation}: We show that in many cases our designs have
overheads low enough for use in deployment (0--15\%) with a study of a diverse
set of real programs ({\em e.g.}, MySQL, memcached and the PARSEC
benchmarks~\cite{parsec}).  }

\end{itemize}

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/systemoverview.pdf}
\Caption{\label{fig:overview}{\bf System Overview.} Last writer slicing tracks provenance, which augments the core dump and enables provenance debugging.  CTraps uses provenance to convey communication to CTraps applications, which implement concurrency analyses, like communication graphs~\cite{recon} or CBI~\cite{cci}.}
\end{figure}

>>>>

\newcount\draft\draft=0 % set to 0 for submission or publication
\newcount\cameraready\cameraready=1

\documentclass[10pt,nocopyrightspace]{sigplanconf}

\ifnum\draft=1
  \input{revision}
  \usepackage{drafthead}
\fi
\usepackage{xxx}
\usepackage{url}
\usepackage{flushend}
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\graphicspath{{./figs/} {./plots/}}

\usepackage[normalem]{ulem}
\usepackage{fancyhdr}

\let\OLDthebibliography\thebibliography
\renewcommand\thebibliography[1]{
  \OLDthebibliography{#1}
  \setlength{\parskip}{0pt}
  \setlength{\itemsep}{0pt plus 0.3ex}
}

@@

\usepackage{xspace}
\newcommand{\addtodo}[1]{\textcolor{red}{[To do: #1]}}
\newcommand{\ctrap}{CTrap\xspace}
\newcommand{\ctraps}{CTraps\xspace}
\newcommand{\ctrapsfull}{CTraps-Full\xspace}
\newcommand{\ctrapsmm}{CTraps-NRR\xspace}
\newcommand{\lws}{LWS\xspace}
\newcommand{\lwsfull}{Last Writer Slicing\xspace}
\newcommand{\lwt}{LWT\xspace}
\newcommand{\lwtfull}{Last Writer Table\xspace}

\newcommand{\Caption}[1]{\begin{minipage}{.95\columnwidth} \caption{#1} \end{minipage} \vspace{-1.2ex}}
\newcommand{\CaptionWide}[1]{\caption{#1}\vspace{-2ex}}
%\doublespacing
\authorinfo{Brandon Lucia}{Carnegie Mellon University, Department of Electrical and Computer Engineering}{blucia@ece.cmu.edu}
\authorinfo{Luis Ceze}{University of Washington, Department of Computer Science and Engineering}{luisceze@cs.washington.edu}

%\author{Brandon Lucia\thanks{Work done in part as a PhD student at University of Washington.}\\Carnegie Mellon University\\{\em blucia@cmu.edu} \and Luis Ceze\\University of Washington\\{\em luisceze@cs.washington.edu}}

%{blucia@cmu.edu luisceze@cs.washington.edu\\{\bf In Submission --- Please Do Not Distribute}}

@@

 \thispagestyle{fancy}
  %For papers in which all authors are employed by the US government, un-comment the following
  %\def\ieeenotice{U.S. Government work not protected by U.S. copyright}
  %For papers in which all authors are employed by a Crown government (UK, Canada, and Australia), uncomment the following
  %\def\ieeenotice{978-1-4799-8161-8/15/\$31.00 {\copyright}2015 Crown}
  %For all other papers, uncomment the following:
  \def\ieeenotice{978-1-4799-8161-8/15/\$31.00 {\copyright}2015 IEEE}
  \def\confinfo{2015 IEEE/ACM International Symposium on Code Generation and Optimization}
  \lfoot{\confinfo\\\ieeenotice} \cfoot{}
  \renewcommand{\headrulewidth}{0pt}

@@

\begin{abstract}
%Monitoring inter-thread memory dependences in production is useful.  Doing so
%improves debugging information and provides information to dynamic analyses and
%runtime systems.  Prior work that continuously monitors cross-thread
%dependences typically has overheads too high for production.  
We propose {\em \lwsfull} (\lws), a mechanism for tracking {\em data
provenance information} in multithreaded code in a production setting.  Last
writer slices dynamically track provenance of values by recording the thread
and operation that last wrote each variable. We show that this information
complements core dumps and greatly improves debugability. We also propose {\em
communication traps} (\ctraps), an extensible framework that uses \lws to interpose on operations that lead to inter-thread communication. We
show how to use \ctraps to implement multiple dynamic analysis tools for
mulithreaded programs. Our evaluation on server programs and PARSEC benchmarks
shows that \lws has low run time overheads (0--15\%) for many
applications,including  memcached, LevelDB, MySQL, and Apache. Our debugging
case studies of real, buggy programs shows that \lws are indeed
useful for debugging.  
\end{abstract}

@@

Our approach is to provide the programmer with data {\em provenance}
information that can be examined alongside the program's memory state ({\em
i.e.}, core dump).  Provenance information describes why a memory location
contains its value.  In this work we develop a new provenance tracking
mechanism called {\em \lwsfull} (\lws).  A memory location's last writer slice is a
record containing the program point and the identifier of the thread that last
wrote data to that memory location.  As the program executes in production, we
collect last writer slices for each potentially shared memory location,
recording the provenance of the data in each location.  Last writer slices are
saved with a core dump and are available in the debugger.

@@

Motivated by these {\em ad hoc} efforts, we develop a dynamic analysis
framework called {\em Communication Traps} or {\em \ctraps}.  \ctraps uses the
provenance information provided by last writer slices to identify memory
accesses that lead to {\em communication} between the threads. Communication
occurs when a thread accesses memory last written by a different thread and
\ctraps delivers a software {\em trap} to a thread when it communicates with
another thread, giving it information about how communication occurred.
Programmer-defined {\em \ctraps Applications} can implement arbitrary analysis
using that information. By default, \ctraps has low overheads, appropriate for
production in many cases.  As our evaluation shows, the overhead of \ctraps
applications scales with the complexity of their analysis.
Figure~\ref{fig:overview} shows the relationship between last writer
slices, \ctraps, and \ctraps applications.

@@

\paragraph{Contributions of This Work}
To summarize, this work makes several contributions:

\noindent {\bf Last writer slicing} ({\em LWS}): a provenance tracking mechanism
useful for concurrency debugging.  We present an \lws implementation
efficient enough for production use.

\vspace{.5ex}
\noindent {\bf Communication traps} (\ctraps): an extensible
framework for implementing concurrency analyses, built on \lws.

\vspace{.5ex}
\noindent {\bf Efficacy evaluation}: We show that \lws is useful for debugging with
a study of several real concurrency bugs.  We show \ctraps is useful for
building concurrency analyses by implementing two analyses from prior
work~\cite{cci,defuse,recon}.

\vspace{.5ex}
\noindent {\bf Performance evaluation}: We show that in many cases our designs have
overheads low enough for use in deployment (0--15\%) with a study of a diverse
set of real programs ({\em e.g.}, MySQL, memcached and the PARSEC
benchmarks~\cite{parsec}).

\begin{figure}[t]
\centering
\includegraphics[width=\columnwidth]{figs/systemoverview.pdf}
\Caption{\label{fig:overview}{\bf System Overview.} Last writer slicing tracks provenance, which augments the core dump and enables provenance debugging.  CTraps uses provenance to convey communication to CTraps applications, which implement concurrency analyses, like communication graphs~\cite{recon} or CBI~\cite{cci}.}
\end{figure}
